Computing astrophysical inferences with diffusive nested sampling

### 2014-03-23 ms.tex
Figure (data)
Caption
Talk about the information

For the ``easy'' simulated data set containing 100 stars, the information
$H$ was approximately 500 nats, whereas for the ``hard'' dataset the information
was $\sim$2600 nats. In both cases, the upward part of the DNS run was
straightforward. However, in the ``hard'' case the diffusion phase of DNS
is very slow. Unfortunately, I am presently unaware of any Nested Sampling
method that only moves upwards in likelihood but is able to correct for the
use of imperfect MCMC. Variations of annealing or tempering such as
``annealed importance sampling'' and ``linked importance sampling'' \citep{ais, lis}
have both these properties, suggesting this may be possible with NS as well.

>>>>

\begin{figure}
%\begin{center}
\includegraphics[scale=0.5]{Figures/data.pdf}
\caption{The two simulated datasets originally studied 
in \citet{starfield}. The ``easy'' image contains 100 stars and the
``hard'' image contains 1000. The goal is to infer the number $N$ of stars
in an image along with their positions and brightnesses.\label{fig:data}}
%\end{center}
\end{figure}

One measure of the scale of an inference problem is the information
$\mathcal{H}$, defined as the KL divergence from the prior to the posterior:
\begin{eqnarray}
\mathcal{H} = \int p(\pars | \data) \log\left[\frac{p(\pars | \data)}{p(\pars)}\right]\, d^n\pars
\end{eqnarray}
An intuitive interpretation of $\mathcal{H}$ is that it measures how small
the posterior volume is relative to the prior. For example, if the prior
were uniform over a volume $V$ and the posterior were uniform over a smaller
volume $V'$, then $\mathcal{H}$ would equal $\log(V/V')$. Nested Sampling allows
$\mathcal{H}$ to be calculated.

For the ``easy'' simulated data set containing 100 stars, the information
was approximately 500 nats, whereas for the ``hard'' dataset the information
was $\sim$2600 nats. In both cases, the upward part of the DNS run was
straightforward. However, in the ``hard'' case the diffusion phase of DNS
is very slow. While the time taken to compress from the prior to the posterior
in the upwards phase of DNS is proportional to $\mathcal{H}$, the time taken to
diffuse from the prior to the posterior in the second phase of DNS is
proportional to $\mathcal{H}^2$.
Unfortunately, I am presently unaware of any Nested Sampling
method that only moves upwards in likelihood but is able to correct for the
use of imperfect MCMC. Variations of annealing or tempering such as
``annealed importance sampling'' and ``linked importance sampling'' \citep{ais, lis}
have both these properties, suggesting this may be possible with NS as well.
